# SQL


## General
  *https://www.w3schools.com/sql/default.asp*
  
### Basic commands

#### SELECT

```{r, echo=T, eval=F}

SELECT * FROM CUSTOMER;

SELECT CUSTOMER_ID,  NAME FROM CUSTOMER;

# Generate a calculated column
SELECT
PRODUCT_ID,
DESCRIPTION,
PRICE AS UNTAXED_PRICE,
round(PRICE * 1.07,2) AS TAXED_PRICE
FROM PRODUCT;

# Text concatenation. Concatenation works with any data type.
SELECT NAME,
CITY || ', ' || STATE AS LOCATION
FROM CUSTOMER;
```

#### WHERE

```{r, echo=T, eval=F}

SELECT * FROM STATION_DATA
WHERE year = 2010;

#Use != or <> to get everything but 2010
SELECT * FROM STATION_DATA
WHERE year != 2010;

#AND, OR, IN statements
SELECT * FROM STATION_DATA
WHERE year >= 2005 AND year <= 2010;

#in
SELECT * FROM STATION_DATA
WHERE MONTH IN (3,6,9,12)

#not in
SELECT * FROM STATION_DATA
WHERE MONTH NOT IN (3,6,9,12)

#modulus operator
#modulus returns the remainder instead of the quotient. A remainder of 0 means there is no remainder at all
SELECT * FROM STATION_DATA
WHERE MONTH %  3 = 0

#using where on text
SELECT * FROM STATION_DATA
WHERE report_code in ('513A63', '1F8A7B')

SELECT * FROM STATION_DATA
WHERE length(report_code) = 6

#wildcards
SELECT * FROM STATION_DATA
WHERE report_code LIKE 'A%'

#B as the first, C as the third letter

SELECT * FROM STATION_DATA
WHERE report_code LIKE 'B_C%'

#Other text functions as INSTR, SUBSTR, REPLACE ETC

SELECT * FROM  station_data
WHERE snow_depth IS NULL;

#Use coalesce to conert NULL to 0, "N/A" etc

SELECT * FROM  station_data
WHERE coalesce(precipitation,0) <= 0.5;

SELECT * FROM station_data
WHERE (rain = 1 AND temperature <= 32)
OR snow_depth > 0;


```



## SQL databases and R

### Accessing a database from R

*Access a database from R.*
*The problem with dplyr is that all operations are conducted in-memory and thus the amount of data you can work with is limited by available memory. The database connection essentially removes that limitation.* 

#### Connecting to a database

*Accessing a temp database by downloading it. Dplyr and dbplyr are used in R to point to the database.*



```{r, echo=T, eval=F}
#https://datacarpentry.org/R-ecology-lesson/05-r-and-databases.html
library("RSQLite")
library("plyr")
library("dbplyr")

# Downloading database for test
dir.create("data_raw", showWarnings = FALSE)
download.file(url = "https://ndownloader.figshare.com/files/2292171",
              destfile = "data_raw/portal_mammals.sqlite", mode = "wb")

# Connect to databse
mammals <- DBI::dbConnect(RSQLite::SQLite(), "data_raw/portal_mammals.sqlite")

# Querying with SQL-syntax vs dplyr syntax

#sql
tbl(mammals, sql("SELECT year, species_id, plot_id FROM surveys"))

#dplyr
surveys <- tbl(mammals, "surveys")
surveys %>%
  select(year, species_id, plot_id)

#Checking head(surveys, n = 10) and nrow(surveys) we see that the  surveys at first glance looks like a data frame but there are some differences.*
head(surveys, n = 10)
nrow(surveys)

#The reason for this is that dplyr dosen´t see the full dataset, only what was asked for when the question in dplyr was translated into SQL.

```

### Running SQL syntax in R


```{r, echo=T, eval=F}
library(sqldf)
#https://dept.stat.lsa.umich.edu/~jerrick/courses/stat701/notes/sql.html#introduction

sqldf('SELECT age, circumference FROM Orange WHERE Tree = 1 ORDER BY circumference ASC')

sqldf("SELECT * FROM iris")

#example
data(BOD)
BOD

#Wildcard: used to extract everything
bod2 <- sqldf('SELECT * FROM BOD')
bod2

#LIMIT controls the number of results
sqldf('SELECT * FROM iris LIMIT 5')

#ORDER BY syntax: ORDER BY var1 {ASC/DESC}, var2 {ASC/DESC}
sqldf("SELECT * FROM Orange ORDER BY age ASC, circumference DESC LIMIT 5")

#Where can be used to add conditional statements
sqldf('SELECT demand FROM BOD WHERE Time < 3')

#WHERE with AND and OR
sqldf('SELECT * FROM rock WHERE (peri > 5000 AND shape < .05) OR perm > 1000')

#IN is used to similiar to %in%
sqldf('SELECT * FROM BOD WHERE Time IN (1,7)')
sqldf('SELECT * FROM BOD WHERE Time NOT IN (1,7)')

#LIKE weak expression command
sqldf('SELECT * FROM chickwts WHERE feed LIKE "%bean" LIMIT 5')
sqldf('SELECT * FROM chickwts WHERE feed NOT LIKE "%bean" LIMIT 5')

#Aggregated data: AVG, MEDIAN, MAX, MIN, SUM
sqldf("SELECT AVG(circumference) FROM Orange")

#SELECT COUNT
d <- data.frame(a = c(1,1,1), b = c(1,NA,NA))
d
sqldf("SELECT COUNT() as numrows FROM d")
sqldf("SELECT COUNT(b) FROM d")



```
