# SQL


## Basic
  *https://www.w3schools.com/sql/default.asp*
  *'Getting started with SQL' using SQLite.*
  *https://github.com/thomasnield/oreilly_getting_started_with_sql*


### SELECT

```{r, echo=T, eval=F}

SELECT * FROM CUSTOMER;

SELECT CUSTOMER_ID,  NAME FROM CUSTOMER;

# Generate a calculated column
SELECT
PRODUCT_ID,
DESCRIPTION,
PRICE AS UNTAXED_PRICE,
round(PRICE * 1.07,2) AS TAXED_PRICE
FROM PRODUCT;

# Text concatenation. Concatenation works with any data type.
SELECT NAME,
CITY || ', ' || STATE AS LOCATION
FROM CUSTOMER;
```

#### WHERE

```{r, echo=T, eval=F}

SELECT * FROM STATION_DATA
WHERE year = 2010;

#Use != or <> to get everything but 2010
SELECT * FROM STATION_DATA
WHERE year != 2010;

#AND, OR, IN statements
SELECT * FROM STATION_DATA
WHERE year >= 2005 AND year <= 2010;

#in
SELECT * FROM STATION_DATA
WHERE MONTH IN (3,6,9,12)

#not in
SELECT * FROM STATION_DATA
WHERE MONTH NOT IN (3,6,9,12)

#modulus operator
#modulus returns the remainder instead of the quotient. A remainder of 0 means there is no remainder at all
SELECT * FROM STATION_DATA
WHERE MONTH %  3 = 0

#using where on text
SELECT * FROM STATION_DATA
WHERE report_code in ('513A63', '1F8A7B')

SELECT * FROM STATION_DATA
WHERE length(report_code) = 6

#wildcards
SELECT * FROM STATION_DATA
WHERE report_code LIKE 'A%'

#B as the first, C as the third letter

SELECT * FROM STATION_DATA
WHERE report_code LIKE 'B_C%'

#Other text functions as INSTR, SUBSTR, REPLACE ETC

SELECT * FROM  station_data
WHERE snow_depth IS NULL;

#Use coalesce to conert NULL to 0, "N/A" etc

SELECT * FROM  station_data
WHERE coalesce(precipitation,0) <= 0.5;

SELECT * FROM station_data
WHERE (rain = 1 AND temperature <= 32)
OR snow_depth > 0;


```

### GROUP BY

*WHERE filters individual records while HAVING filters aggregations.*
```{r, echo=T, eval=F}

SELECT year, month, COUNT(*) AS record_count FROM station_data
WHERE tornado = 1
GROUP BY year, month
ORDER BY year, month

#aggregate
SELECT month, AVG(temperature) as avg_temp
FROM station_data
WHERE year >= 2000
GROUP BY month

#HAVING statement
#To be able to filter on aggregate data you need to use HAVING instead of WHERE

SELECT year,
SUM(precipitation) as total_precipitation
FROM station_data
GROUP BY year
HAVING sum(precipitation) > 30



```


### CASE statements

*CASE statements are read from top to bottom, so the first true condition is the one it uses.*
*A great benefit of CASE statements compared to WHERE is that you can apply different filters for different aggregate values.*
```{r, echo=T, eval=F}

SELECT report_code, year, month, day, wind_speed,

CASE
 WHEN wind_speed >= 40 THEN 'HIGH'
 WHEN wind_speed >= 30 THEN 'MODERATE'
 ELSE 'LOW'
 END as wind_severity
 
FROM station_data

# Use of CASE to apply different filters
SELECT year, month,

SUM(CASE WHEN tornado = 1 THEN precipitation ELSE 0 END) as tornado_precipitation,

SUM(CASE WHEN tornado = 0 THEN precipitation ELSE 0 END) as non_tornado_precipitation

FROM station_data
GROUP BY year, month

# Use of boolean expression

SELECT month, 

AVG(CASE WHEN rain OR hail THEN temperature ELSE null END) AS avg_precipitation_temp,

AVG(CASE WHEN NOT (rain OR hail) THEN temperature ELSE null END) AS avg_non_precipitation_temp

FROM station_data
WHERE year > 2000
GROUP BY month

```


### JOIN

*In multiple joins, it may be erroundous to mix LEFT JOIN with INNER JOIN. This is becasue null values cannot be inner joined on and will get filtered out. LEFT JOIN tolerates null values.*

```{r, echo=T, eval=F}
#INNER JOIN
SELECT ORDER_ID,
CUSTOMER.CUSTOMER_ID,
ORDER_DATE,
ORDER_QTY

FROM CUSTOMER INNER JOIN CUSTOMER_ORDER
ON CUSTOMER.CUSTOMER_ID = CUSTOMER_ORDER.CUSTOMER_ID


#LEFT JOIN

SELECT ORDER_ID,
CUSTOMER.CUSTOMER_ID,
ORDER_DATE,
ORDER_QTY

FROM CUSTOMER LEFT JOIN CUSTOMER_ORDER
ON CUSTOMER.CUSTOMER_ID = CUSTOMER_ORDER.CUSTOMER_ID

#Checking for NULL values
SELECT CUSTOMER.CUSTOMER_ID,
NAME AS CUSTOMER_NAME

FROM CUSTOMER LEFT JOIN CUSTOMER_ORDER
ON CUSTOMER.CUSTOMER_ID = CUSTOMER_ORDER.CUSTOMER_ID

WHERE ORDER_ID IS NULL

#Multiple joins
SELECT
ORDER_ID,
CUSTOMER.CUSTOMER_ID,
'NAME AS CUSTOMER.NAME,'
STREET_ADDRESS,
CITY,
STATE,
ZIP,
ORDER_DATE,
PRODUCT_ID,
DESCRIPTION,
ORDER_QTY

FROM CUSTOMER

INNER JOIN 
CUSTOMER_ORDER ON CUSTOMER_ORDER.CUSTOMER_ID = CUSTOMER.CUSTOMER_ID

INNER JOIN 
PRODUCT ON CUSTOMER_ORDER.PRODUCT_ID = PRODUCT.PRODUCT_ID

#Use coalesce() to turn nulls into zeros.

SELECT
CUSTOMER.CUSTOMER_ID,
NAME AS CUSTOMER_NAME,
coalesce(sum(ORDER_QTY * PRICE), 0) as TOTAL_REVENUE

FROM CUSTOMER

LEFT JOIN CUSTOMER_ORDER
ON CUSTOMER.CUSTOMER_ID = CUSTOMER_ORDER.CUSTOMER_ID

LEFT JOIN PRODUCT
ON CUSTOMER_ORDER.PRODUCT_ID = PRODUCT.PRODUCT_ID

GROUP BY 1,2

```



## SQL databases and R


### Connecting to a database from R

*Access a database from R.*
*The problem with dplyr is that all operations are conducted in-memory and thus the amount of data you can work with is limited by available memory. The database connection essentially removes that limitation.* 
*Accessing a temp database by downloading it. Dplyr and dbplyr are used in R to point to the database.*



```{r, echo=T, eval=F}
#https://datacarpentry.org/R-ecology-lesson/05-r-and-databases.html
library("RSQLite")
library("plyr")
library("dbplyr")

# Downloading database for test
dir.create("data_raw", showWarnings = FALSE)
download.file(url = "https://ndownloader.figshare.com/files/2292171",
              destfile = "data_raw/portal_mammals.sqlite", mode = "wb")

# Connect to databse
mammals <- DBI::dbConnect(RSQLite::SQLite(), "data_raw/portal_mammals.sqlite")

# Querying with SQL-syntax vs dplyr syntax

#sql
tbl(mammals, sql("SELECT year, species_id, plot_id FROM surveys"))

#dplyr
surveys <- tbl(mammals, "surveys")
surveys %>%
  select(year, species_id, plot_id)

#Checking head(surveys, n = 10) and nrow(surveys) we see that the  surveys at first glance looks like a data frame but there are some differences.*
head(surveys, n = 10)
nrow(surveys)

#The reason for this is that dplyr dosen´t see the full dataset, only what was asked for when the question in dplyr was translated into SQL.

```

### Running SQL syntax in R


```{r, echo=T, eval=F}
library(sqldf)
#https://dept.stat.lsa.umich.edu/~jerrick/courses/stat701/notes/sql.html#introduction

sqldf('SELECT age, circumference FROM Orange WHERE Tree = 1 ORDER BY circumference ASC')

sqldf("SELECT * FROM iris")

#example
data(BOD)
BOD

#Wildcard: used to extract everything
bod2 <- sqldf('SELECT * FROM BOD')
bod2

#LIMIT controls the number of results
sqldf('SELECT * FROM iris LIMIT 5')

#ORDER BY syntax: ORDER BY var1 {ASC/DESC}, var2 {ASC/DESC}
sqldf("SELECT * FROM Orange ORDER BY age ASC, circumference DESC LIMIT 5")

#Where can be used to add conditional statements
sqldf('SELECT demand FROM BOD WHERE Time < 3')

#WHERE with AND and OR
sqldf('SELECT * FROM rock WHERE (peri > 5000 AND shape < .05) OR perm > 1000')

#IN is used to similiar to %in%
sqldf('SELECT * FROM BOD WHERE Time IN (1,7)')
sqldf('SELECT * FROM BOD WHERE Time NOT IN (1,7)')

#LIKE weak expression command
sqldf('SELECT * FROM chickwts WHERE feed LIKE "%bean" LIMIT 5')
sqldf('SELECT * FROM chickwts WHERE feed NOT LIKE "%bean" LIMIT 5')

#Aggregated data: AVG, MEDIAN, MAX, MIN, SUM
sqldf("SELECT AVG(circumference) FROM Orange")

#SELECT COUNT
d <- data.frame(a = c(1,1,1), b = c(1,NA,NA))
d
sqldf("SELECT COUNT() as numrows FROM d")
sqldf("SELECT COUNT(b) FROM d")



```
