# HTML, CSS and JAVASCRIPT


## HTML
  
### Bond Calculator in R

#### Calculate Forward Rates

**Calculate Forward Rates**
  

```{r, echo=T, eval=F}

``` 

## CSS
  
### Bond Calculator in R

#### Calculate Forward Rates

**Calculate Forward Rates**
  

```{r, echo=T, eval=F}

``` 

## JAVASCRIPT

### Operators

| Operator    | Description               | Example | 
| :----------- | :----------------------: | -----------: |
| ??      | Returns true if both operands evaluate to true | true && false (false) |
| ||      | Returns true if either operand evaluates to true, otherwise false | if  (x == y || y > z) |
| !       | Inverts the boolean value of the operand | !true (false) |

### Statements

#### If statements
  

```{r, echo=T, eval=F}

# Multiple statements
if(condition) {
  statement_1;
  statement_2;
  statement_3;
 }
 
 # Else clause
 if(condition) {
  statement(s);
  } else {
  statement(s);
 }
 
 #Else if clause. Branching logical structure can get very complex.
 if(condition) {
  statement(s);
 } else if(condition) {
  statement(s);
 }
 
 #Switch statement
 switch(condition) {
  case label_1:
    statement(s);
    break;
   
   case label_2:
    statement(s);
    break;
   
   default:
    statement(s);
    break;
}

#While loops
while(condition) {
  statement(s);
}

#Do while
do {
  statement(s);
} while(condition);

#For loops
for(initialize ; condition ; iteration) {
  statement(s);
}
``` 
