# HTML, CSS and JAVASCRIPT


## HTML
  
### Basic

#### Basic

**Basic**
  

```{r, echo=T, eval=F}

``` 

## CSS
  
### Basic

#### Basic

*Test*
  

```{r, echo=T, eval=F}

``` 

## JAVASCRIPT

### Operators

| Operator    | Description               | Example | 
| :----------- | :----------------------: | -----------: |
| ??      | Returns true if both operands evaluate to true | true && false (false) |
| ||      | Returns true if either operand evaluates to true, otherwise false | if  (x == y || y > z) |
| !       | Inverts the boolean value of the operand | !true (false) |



### Anatomy

#### Keywords

**JavaScript defines a set of keywords known as reserved words, these include: break, case, catch, continue, default, delete, do, else, false, finally, for, function, if, in, instanceof, new, null, return, switch, this, throw, true, try, typeof, var, void, while, with.**


| Operator    | Description               | Example | 
| :----------- | :----------------------: | -----------: |
| This      | Returns true if both operands evaluate to true | true && false (false) |
| ||      | Returns true if either operand evaluates to true, otherwise false | if  (x == y || y > z) |
| !       | Inverts the boolean value of the operand | !true (false) |


#### If statements
  
```{r, echo=T, eval=F}

# Multiple statements
if(condition) {
  statement_1;
  statement_2;
  statement_3;
 }
 
 # Else clause
 if(condition) {
  statement(s);
  } else {
  statement(s);
 }
 
 #Else if clause. Branching logical structure can get very complex.
 if(condition) {
  statement(s);
 } else if(condition) {
  statement(s);
 }
 
 #Switch statement
 switch(condition) {
  case label_1:
    statement(s);
    break;
   
   case label_2:
    statement(s);
    break;
   
   default:
    statement(s);
    break;
}
``` 



#### Loops
  
```{r, echo=T, eval=F}
#While loops
while(condition) {
  statement(s);
}

#Do while
do {
  statement(s);
} while(condition);

#For loops
for(initialize ; condition ; iteration) {
  statement(s);
}
``` 

### Functions and classes

#### Named and anonymous functions

*Using anonymous functions we use a variable name instead of a function name.*
```{r, echo=T, eval=F}
#//Named function
function calculateRectArea(rectWidth, rectHeight) {
  return(rectWidth * rectheight)
}

#//Anonymous function
var calculateRectHypotenuse = function(rectWidth, rectHeight) {
  var widthsquared = rectWidth * rectWidth;
  var heigthSquared = rectHeight * rectHeight;
  return(Math.sqrt(widthSquared + heightSquared));
 }
 
#//calling
var myRectArea = calculateRectArea(49, 28);
var myRectHypotenuse = calculateRectHypotenuse(49,28);

``` 

#### Namespace

*Namespaces are notional spaces or contexts in which a set of names can be defined, to avoid name clashed. Two names can be identical as long as they exist in different namespace.

```{r, echo=T, eval=F}
#//Creating a namespace in js script. It is a good practise to include your namespace declaration in top of all your scripts. Functions needs to be anonymous.
#//First we create an object called com inside the global namespace, and then we create an object called flametreepublishing inside the com project.
#// After declaring, but not initializing, we test that each namespace object does not exist before using it.


var com;
if(!com) {
	com = {};
}

if (!com.flametreepublishing) {
	com.flametreepublishing = {}
}


```

#### Object-Oriented

*While JavaScript is not a true OOP language, it is still used in a OOP style. We write scripts called classes that define properties.*

```{r, echo=T, eval=F}
#//Example: Define a custom class, then create an instance.

var com;
if(!com) {
	com = {};
}
if(!com.flametreepublishing) {
	com.flametreepublishing = {};
}

com.flametreepublishing.QuizQuestion = functio(aQustionNum, AQuestionText, aAnswer, aCorrectAnswerIndex){
	this.questionNum = aQustionNum;
	this.questionText = AQuestionText;
	this.answers = aAnswer;
	this.correctAnswerIndex = aCorrectAnswerIndex;


#Creating an instance
new com.flametreepublishing.QuizQuestion(
	1,
	"Approx how far away from the Earth is the Sun?",
	["200 miles", "93 000 000 miles", "49 000 000 miles", "150 000 miles"],
	1
)
}

```


### DOM and events of JS

#### DOM programming
*Document Object Model programming allows JavaScript to make changes to a page after it has been loaded.*

```{r, echo=T, eval=F}


#//We've updated substantially the renderQuestion method. Now, rather than
#//relying on document.write, which only works when the page is being loaded,
#//we're now using DOM programming techniques to create HTML elements 
#//on-the-fly from within JavaScript.

com.flametreepublishing.QuizQuestion.prototype.renderQuestion = function() {
#	//First we create a <div> element in which to store the question's
#	//content. This <div> is created but does not yet exist on the page
	var questionDiv = document.createElement("div");
#	//We'll set an id attribute on the <div> - later, this will help us to
#	//identify which question a user clicks on. We'll give the id a leading
#	//'q' because it is bad practice to start an id value with a number.
	questionDiv.id = "q" + this.questionNum;
#	//Now we create an <h2> element for the question's title
	var questionHeading = document.createElement("h2");
#	//An element's innerHtml property allows us to write HTML that will be
#	//rendered within that element...
	questionHeading.innerHTML = "QUESTION " + this.questionNum;
#	//Now we add the <h2> to the <div>
	questionDiv.appendChild(questionHeading);
#	//Next, we create a <p> to hold the question text itself, and add
#	//this to the <div> too.
	var questionTextPara = document.createElement("p");
	questionTextPara.innerHTML = this.questionText;
	questionDiv.appendChild(questionTextPara);
#	//Now we'll loop through the QuizQuestion object's 'answers'
#	//array, creating a <p> for each and and adding them 
#	//to our <div> element
	for(var i = 0; i < this.answers.length; i++) {
		var answerPara = document.createElement("p");
		answerPara.innerHTML = this.answers[i];
		answerPara.id = "a" + i;
		questionDiv.appendChild(answerPara);		
	}
#	//Finally, we add the <div> to the body of the page
	document.body.appendChild(questionDiv);	
}
```
#### Events

**Event Handler**

*W3C event model*


```{r, echo=T, eval=F}

#//Create a method for handling user 'click' events
com.flametreepublishing.SimpleQuiz.prototype.clickHandler = function(e) {
#	//The handler will always be passed an objec that contains data
#	//about the event that triggered the handler. We're using an identifier
#	//of 'e' for this. With a 'click' event, 'e.target' will always refer to 
#	//the HTML element on which the click event occurred.
#	//First we'll get the id attribute of the clicked answer
	var clickedAnswerId = e.target.id;
#	//Now we need to extract the answer index, a Number, from the id value,
#	//which is a string. We do this by extracting the second character of the
#	//id using the 'substr' method of the String class, and then casting this to
#	//a number. Notice that we can do this all within one compound expression.
	var clickedAnswerIndex = Number(clickedAnswerId.substr(1, 1));
#	//Next we need to know which question has been answered. Recall that we added
#	//an id attribute to the <div> that contains the question - this <div> is the
#	//parent of the answer <p> that was clicked, so we can access it using
#	//'e.target.parentNode'.
#	//Once we have a reference to the <div> we can extract the question index in
#	//much the same way as we did the answer index
	var clickedQuestionId = e.target.parentNode.id;
	var clickedQuestionNum = Number(clickedQuestionId.substr(1, 1));
#	//The question number stored in the containg <div> id attribute is one-based, but
#	//we need a zero-based number when accessing the questions array - for this reason
#	//we subtract '1' from the clickedQuestionNum when retrieving the QuizQuestion object
	var clickedQuestion = com.flametreepublishing.simpleQuiz.questions[clickedQuestionNum -1];
#	//Now that we have the correct QuizQuestion object we can call its 'checkUserAnswer'
#	//method to see if the user was correct. Recall that 'checkUserAnswer' returns
#	//'true' or 'false' - all we need do, then, is call the method as the conditional 
#	//expression of an 'if' statement.
	if(clickedQuestion.checkUserAnswer(clickedAnswerIndex)) {
		alert("Correct! Well done.");
	} else {
		alert("No - that's not correct. Try again.");
	}	
}
	
```

**Event Listener**
